Finding Groups Within the Data

### 2016-12-28 articles/avoid_pie_charts/avoid_pie_charts.Rmd
re-run charts with latest ggplot2

---
title: "Faceted Bar Charts"
date: "`r Sys.Date()`"
author: "Ming-Yu Liu"
output:
  rmdformats::readthedown:
    highlight: haddock
---

@@

@@

- geom_bar's `size` controls the bar's border size.
- `expand` is a numeric vector of length two giving multiplicative and additive expansion constants. These constants ensure that the data is placed some distance away from the axes.
- `strip .background / .text` controls the title section for each facet.
- `panel.grid .minor / .major`  controls the grid lines in the plot.
- `panel.margin` contols the margin between facets.
- `as_labeller` the function that changes the strip's label without doesn't changing the underlying data. All you have to do is create a named character vector with the new label mapping to the original label name and call the function and pass it in to `facet_grid`'s `labeller` argument.

@@

ggplot( d, aes( x = Hours, y = Percentage / 100, fill = Hours ) ) +
geom_bar( stat = "identity", width = 0.75, color = "#2b2b2b", size = 0.05 ) + 
scale_y_continuous( labels = percent, limits = c( 0, 0.5 ) ) + 
scale_x_discrete( expand = c( 0, 1 ), labels = x_labels ) + 
scale_fill_manual( values = c( "#a6cdd9", "#d2e4ee", "#b7b079", "#efc750" ) ) +
facet_wrap( ~ Task, labeller = as_labeller(label_names) ) + 
labs( x = NULL, y = NULL, title = "Where Does the Time Go?" ) +
theme( strip.text = element_text( size = 12, color = "white", hjust = 0.5 ),
	   strip.background = element_rect( fill = "#858585", color = NA ),	   
	   panel.background = element_rect( fill = "#efefef", color = NA ),
	   panel.grid.major.x = element_blank(),
	   panel.grid.minor.x = element_blank(),
	   panel.grid.minor.y = element_blank(),
	   panel.grid.major.y = element_line( color = "#b2b2b2" ),
	   panel.margin.x = unit( 1, "cm" ),
	   panel.margin.y = unit( 0.5, "cm" ),
	   legend.position = "none" )

@@

- [Worst Chart In The World](http://www.businessinsider.com/pie-charts-are-the-worst-2013-6)
- [How to replace a pie chart](http://varianceexplained.org/r/improving-pie-chart/)
- [Take the Pies out of Pi Day](http://rud.is/b/2016/03/14/spinning-cycles-in-box-4-to-take-the-pies-out-of-pi-day/)

>>>>

---
title: "Faceted Bar Charts"
date: "`r Sys.Date()`"
author: "Ethen Liu"
output:
  rmdformats::readthedown:
    highlight: pygments
---

@@

# Pie Chart

@@

- geom_bar's `size` controls the bar's border size.
- `expand` is a numeric vector of length two giving multiplicative and additive expansion constants. These constants ensure that the data is placed some distance away from the axes.
- `strip .background / .text` controls the title section for each facet.
- `panel.grid .minor / .major`  controls the grid lines in the plot.
- `panel.spacing` contols the margin between facets.
- `as_labeller` the function that changes the strip's label without doesn't changing the underlying data. All you have to do is create a named character vector with the new label mapping to the original label name and call the function and pass it in to `facet_grid`'s `labeller` argument.

@@

ggplot( d, aes( x = Hours, y = Percentage / 100, fill = Hours ) ) +
geom_bar( stat = "identity", width = 0.75, color = "#2b2b2b", size = 0.05 ) + 
scale_y_continuous( labels = percent, limits = c( 0, 0.5 ) ) + 
scale_x_discrete( expand = c( 0, 1 ), labels = x_labels ) + 
scale_fill_manual( values = c( "#a6cdd9", "#d2e4ee", "#b7b079", "#efc750" ) ) +
facet_wrap( ~ Task, labeller = as_labeller(label_names) ) + 
labs( x = NULL, y = NULL, title = "Where Does the Time Go?" ) +
theme( strip.text = element_text( size = 12, color = "white", hjust = 0.5 ),
	   strip.background = element_rect( fill = "#858585", color = NA ),	   
	   panel.background = element_rect( fill = "#efefef", color = NA ),
	   panel.grid.major.x = element_blank(),
	   panel.grid.minor.x = element_blank(),
	   panel.grid.minor.y = element_blank(),
	   panel.grid.major.y = element_line( color = "#b2b2b2" ),
	   panel.spacing.x = unit( 1, "cm" ),
	   panel.spacing.y = unit( 0.5, "cm" ),
	   legend.position = "none" )

@@

- [Blog: Worst Chart In The World](http://www.businessinsider.com/pie-charts-are-the-worst-2013-6)
- [Blog: How to replace a pie chart](http://varianceexplained.org/r/improving-pie-chart/)
- [Blog: Take the Pies out of Pi Day](http://rud.is/b/2016/03/14/spinning-cycles-in-box-4-to-take-the-pies-out-of-pi-day/)

### 2016-12-28 articles/calendar_heatmaps/calendar_heatmaps.Rmd
re-run charts with latest ggplot2

---
title: "Calendar Heatmaps"
author: "Ming-Yu Liu"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: pygments
---

@@

attacks <- fread("data/eventlog.csv")
head(attacks)

@@

make_hr_wkday <- function( ts, sc, tz ){
	# convert each time with the appropriate timezone,
	# the timezone parameter, tz, only takes a single value,
	# then extract its weekdays and hour
	real_times <- ymd_hms( ts, tz = tz[1], quiet = TRUE )
	dt <- data.table( source_country = sc,
					  wkday = weekdays(real_times),
					  hour = hour(real_times) )
	return(dt)
}

# simply stores the name of the weekday
# convert weekday and hour into factor so they'll be ordered
# when plotting
weekday_levels <- levels( weekdays( 0, abbreviate = FALSE ) )
attacks <- attacks %>% 
		   group_by(tz) %>%
		   do( make_hr_wkday( .$timestamp, .$source_country, .$tz ) ) %>% 
		   ungroup() %>% 
		   mutate( wkday = factor( wkday, levels = weekday_levels ),
		   		   hour  = factor( hour, levels = 0:23 ) )
head(attacks)

@@

grouped <- attacks %>% count( wkday, hour ) %>% ungroup()

ggplot( grouped, aes( hour, wkday, fill = n ) ) + 
geom_tile( color = "white", size = 0.1 ) + 
theme_tufte( base_family = "Helvetica" ) + 
coord_equal() + 
scale_fill_viridis( name = "# of Events", label = comma ) + 
labs( x = NULL, y = NULL, title = "Events per weekday & time of day" ) +
theme( axis.ticks = element_blank(),
	   plot.title = element_text( hjust = 0.5 ),
	   legend.title = element_text( size = 8 ),
	   legend.text = element_text( size = 6 ) )

@@

gg <- ggplot( top_country_attacks, aes( x = hour, y = wkday, fill = n ) ) + 
	  geom_tile( color = "white", size = 0.1 ) +
	  scale_fill_viridis( name = "# Events" ) + 
	  coord_equal() + 
	  facet_wrap( ~source_country, ncol = 2 ) +
	  labs( x = NULL, y = NULL, title = "Events per weekday & time of day by country\n" ) + 
	  theme_tufte( base_family = "Helvetica" ) + 
	  theme( axis.ticks = element_blank(),
			 axis.text = element_text( size = 8 ),
			 panel.border = element_blank(),
			 plot.title = element_text( hjust = 0.5 ),
			 strip.text = element_text( hjust = 0.5 ),
			 panel.margin = unit( 0.1, "cm" ),
	  		 legend.position = "bottom",
	  		 legend.title = element_text( size = 8 ),
	  		 legend.text = element_text( size = 6 ),
	  		 legend.key.size = unit( 0.4, "cm" ),
	  		 legend.key.width = unit( 1, "cm" ) )
gg

@@

plots <- lapply( top_country, function(x){

	subset_data <- top_country_attacks %>% filter( source_country == x )
	gg <- ggplot( subset_data, aes( x = hour, y = wkday, fill = n ) ) + 
		  geom_tile( color = "white", size = 0.1 ) +
		  scale_fill_viridis( name = "# Events" ) + 
		  scale_y_discrete( expand = c( 0, 0 ) ) +
		  coord_equal() + 
		  labs( x = NULL, y = NULL, title = x ) + 
		  theme_tufte( base_family = "Helvetica" ) + 
		  theme( axis.ticks = element_blank(),
				 axis.text = element_text( size = 7 ),
				 panel.border = element_blank(),
				 plot.title = element_text( hjust = 0.5 ),
				 strip.text = element_text( hjust = 0.5 ),
				 panel.margin = unit( 0.1, "cm" ),
		  		 legend.position = "bottom",
		  		 legend.title = element_text( size = 6 ),
		  		 legend.text = element_text( size = 6 ),
		  		 legend.key.size = unit( 0.4, "cm" ) )
	return(gg)
})

@@

- [Making faceted heatmaps with ggplot2](https://rud.is/b/2016/02/14/making-faceted-heatmaps-with-ggplot2/)
- [Big Data Insights - IT Support Log Analysis](http://costaleconomist.blogspot.tw/2015/12/big-data-insights-it-support-log.html)

>>>>

---
title: "Calendar Heatmaps"
author: "Ethen Liu"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: pygments
---

@@

attacks <- fread("data/eventlog.csv")
knitr::kable( head(attacks) )

@@

make_hr_wkday <- function(ts, sc, tz) {
	# convert each time with the appropriate timezone,
	# the timezone parameter, tz, only takes a single value,
	# then extract its weekdays and hour
	real_times <- ymd_hms( ts, tz = tz[1], quiet = TRUE )
	dt <- data.table( source_country = sc,
					  wkday = weekdays(real_times),
					  hour = hour(real_times) )
	return(dt)
}

# convert weekday and hour into factor so they'll be ordered when plotting
wkday_levels <- c('Sunday', 'Monday', 'Tuesday', 'Wednesday', 
				  'Thursday', 'Friday', 'Saturday')
attacks <- attacks %>% 
		   group_by(tz) %>%
		   do( make_hr_wkday( .$timestamp, .$source_country, .$tz ) ) %>% 
		   ungroup() %>% 
		   mutate( wkday = factor(wkday, levels = wkday_levels),
		   		   hour  = factor(hour, levels = 0:23) )

knitr::kable( head(attacks) )

@@

grouped <- attacks %>% count(wkday, hour) %>% ungroup()

ggplot( grouped, aes(hour, wkday, fill = n) ) + 
geom_tile(color = "white", size = 0.1) + 
theme_tufte(base_family = "Helvetica") + 
coord_equal() + 
scale_fill_viridis(name = "# of Events", label = comma) + 
labs(x = NULL, y = NULL, title = "Events per weekday & time of day") +
theme( axis.ticks = element_blank(),
	   plot.title = element_text(hjust = 0.5),
	   legend.title = element_text(size = 8),
	   legend.text = element_text(size = 6) )

@@

gg <- ggplot( top_country_attacks, aes(x = hour, y = wkday, fill = n) ) + 
	  geom_tile(color = "white", size = 0.1) +
	  scale_fill_viridis(name = "# Events") + 
	  coord_equal() + 
	  facet_wrap( ~source_country, ncol = 2 ) +
	  labs(x = NULL, y = NULL, title = "Events per weekday & time of day by country\n") + 
	  theme_tufte(base_family = "Helvetica") + 
	  theme( axis.ticks = element_blank(),
			 axis.text = element_text(size = 8),
			 panel.border = element_blank(),
			 plot.title = element_text(hjust = 0.5),
			 strip.text = element_text(hjust = 0.5),
			 panel.margin = unit(0.1, "cm"),
	  		 legend.position = "bottom",
	  		 legend.title = element_text(size = 8),
	  		 legend.text = element_text(size = 6),
	  		 legend.key.size = unit(0.4, "cm"),
	  		 legend.key.width = unit(1, "cm") )
gg

@@

plots <- lapply(top_country, function(x) {

	subset_data <- top_country_attacks %>% filter(source_country == x)
	gg <- ggplot( subset_data, aes(x = hour, y = wkday, fill = n) ) + 
		  geom_tile(color = "white", size = 0.1) +
		  scale_fill_viridis(name = "# Events") + 
		  scale_y_discrete( expand = c(0, 0) ) +
		  coord_equal() + 
		  labs(x = NULL, y = NULL, title = x) + 
		  theme_tufte(base_family = "Helvetica") + 
		  theme( axis.ticks = element_blank(),
				 axis.text = element_text(size = 7),
				 panel.border = element_blank(),
				 plot.title = element_text(hjust = 0.5),
				 strip.text = element_text(hjust = 0.5),
				 panel.margin = unit(0.1, "cm"),
		  		 legend.position = "bottom",
		  		 legend.title = element_text(size = 6),
		  		 legend.text = element_text(size = 6),
		  		 legend.key.size = unit(0.4, "cm") )
	return(gg)
})

@@

- [Blog: Making faceted heatmaps with ggplot2](https://rud.is/b/2016/02/14/making-faceted-heatmaps-with-ggplot2/)
- [Blog: Big Data Insights - IT Support Log Analysis](http://costaleconomist.blogspot.tw/2015/12/big-data-insights-it-support-log.html)

### 2016-12-28 articles/nyt_scatter/nyt_scatter.Rmd
re-run charts with latest ggplot2

---
title: "Scatter Plot"
date: "`r Sys.Date()`"
author: "Ming-Yu Liu"
output:
  rmdformats::readthedown:
    highlight: haddock
---

@@

# 2016.5.12 you have to install the development version of ggplot2
# devtools::install_github("hadley/ggplot2") or subtitles won't work
library(tidyr)
library(dplyr)
library(readr)
library(scales)
library(ggplot2) 
setwd("/Users/ethen/Business-Analytics/articles/nyt_scatter")

file <- "nytimes_vote.tsv"
if( !file.exists(file) ){
	url <- "https://static01.nyt.com/newsgraphics/2016/04/21/undervote/ad8bd3e44231c1091e75621b9f27fe31d116999f/data.tsv"
	download.file( url, file )
}

# convert to more informative column name
df <- read_tsv(file)
df <- df %>% rename( "Someone else" = undervt, 
					 "Hilary Clinton" = clintonpct, 
					 "Bernie Sanders" = sanderspct )
df

df <- gather( select( df, -tvotes ), party, pct, -ratio, -fips )
df <- arrange( df, fips, ratio )
df <- df %>% mutate( 
	party = factor( party, levels = c("Hilary Clinton", "Bernie Sanders", "Someone else" ) )
)
df

@@

ggplot( df, aes( x = ratio, y = pct ) ) + 
geom_point( aes( fill = party ), size = 3, alpha = 0.8, color = "white", shape = 21 ) + 
scale_fill_manual( name = "", values = fill_color ) +
theme_bw( base_family = "Arial Narrow" ) + 
scale_y_continuous( label = percent, limits = c( 0, 1.05 ) ) +
scale_x_continuous( limits = c( 0, 4.5 ), breaks = seq( 0, 4.5, 0.5 ) ) +
geom_text( data = data.frame( label = "↑ Share of 2016 primary vote"),
		   aes( x = 0, y = 1, label = label ), vjust = -1, hjust = 0, size = 3,
		   fontface = "bold", family = "Arial Narrow" ) + 
labs( x = "Ratio of registered Democrats to Obama voters →", 
	  y = NULL, title = "The Kinds of Places Sanders Beats Clinton",
	  subtitle = "Each dot on this chart represents the share of a county's vote for a candidate in the 2016 Democratic primary" ) +
theme( legend.key = element_blank(), 
	   legend.position = "top",
	   plot.title = element_text( face = "bold" ),
	   axis.ticks = element_blank(),
	   axis.text = element_text( size = 8 ),
	   axis.title.x = element_text( hjust = 1, face = "bold", size = 9 ),
	   panel.grid.minor = element_blank(),
	   panel.grid.major = element_line( linetype = "dotted", size = 0.5 ),
	   panel.border = element_blank(),
	   plot.margin = margin( t = 10, r = 10, b = 10, l = 10 ) )

@@

- [Gist](https://gist.github.com/hrbrmstr/4632dc9cefc7566b412ec9945a6c4922)

>>>>

---
title: "Scatter Plot"
date: "`r Sys.Date()`"
author: "Ethen Liu"
output:
  rmdformats::readthedown:
    highlight: pygments
---

@@

library(tidyr)
library(dplyr)
library(readr)
library(scales)
library(ggplot2) 
setwd("/Users/ethen/Business-Analytics/articles/nyt_scatter")

file <- "nytimes_vote.tsv"
if( !file.exists(file) ) {
	url <- "https://static01.nyt.com/newsgraphics/2016/04/21/undervote/ad8bd3e44231c1091e75621b9f27fe31d116999f/data.tsv"
	download.file(url, file)
}

# convert to more informative column name
df <- read_tsv(file)
df <- df %>% rename("Someone else" = undervt, 
					"Hilary Clinton" = clintonpct, 
					"Bernie Sanders" = sanderspct)

df <- gather( select(df, -tvotes), party, pct, -ratio, -fips )
df <- arrange(df, fips, ratio)
df <- df %>% mutate( 
	party = factor( party, levels = c("Hilary Clinton", "Bernie Sanders", "Someone else") )
)
df

@@

ggplot( df, aes(x = ratio, y = pct) ) + 
geom_point( aes(fill = party), size = 3, alpha = 0.8, color = "white", shape = 21 ) + 
scale_fill_manual(name = "", values = fill_color) +
theme_bw( base_family = "Arial Narrow" ) + 
scale_y_continuous( label = percent, limits = c(0, 1.05) ) +
scale_x_continuous( limits = c(0, 4.5), breaks = seq(0, 4.5, 0.5) ) +
geom_text( data = data.frame(label = "↑ Share of 2016 primary vote"),
		   aes(x = 0, y = 1, label = label), vjust = -1, hjust = 0, size = 3,
		   fontface = "bold", family = "Arial Narrow" ) + 
labs( x = "Ratio of registered Democrats to Obama voters →", 
	  y = NULL, title = "The Kinds of Places Sanders Beats Clinton",
	  subtitle = "Each dot on this chart represents the share of a county's vote for a candidate in the 2016 Democratic primary" ) +
theme( legend.key = element_blank(), 
	   legend.position = "top",
	   plot.title = element_text(face = "bold"),
	   axis.ticks = element_blank(),
	   axis.text = element_text(size = 8),
	   axis.title.x = element_text(hjust = 1, face = "bold", size = 9),
	   panel.grid.minor = element_blank(),
	   panel.grid.major = element_line(linetype = "dotted", size = 0.5),
	   panel.border = element_blank(),
	   plot.margin = margin(t = 10, r = 10, b = 10, l = 10) )

@@

- [Gist: nytscatter.r](https://gist.github.com/hrbrmstr/4632dc9cefc7566b412ec9945a6c4922)

### 2016-12-28 finding_groups/finding_groups.Rmd
re-run charts with latest ggplot2

---
title: "Finding Groups Within the Data"
author: "Ming-Yu Liu"
date: "2016-04-05"
output:
  rmdformats::readthedown:
    highlight: pygments
---

@@

ggplot( data_sku, aes( CV, ADS ) ) + geom_point() +  
geom_hline( yintercept = 4, color = "blue" ) + 
geom_vline( xintercept = .2, color = "blue" ) + 
labs( x = "Coefficient of Variation", y = "Average Daily Sales", 
	  title = "SKU Example" )

@@

# hierarchical clustering, with k (clustering number) = 3 
d <- dist( scale(data_sku), method = "euclidean" )
cluster <- hclust( d, method = "ward.D" )
data_sku[ ,  groups := as.factor( cutree( cluster, k = 3 ) ) ]
head(data_sku)

@@

ggplot( data_sku, aes( CV, ADS, color = groups ) ) + geom_point() + 
labs( x = "Coefficient of Variation", y = "Average Daily Sales", 
	  title = "SKU Example" )

@@

4. For clustering methods, the number of segmentation is a parameter that has to be user-specified before running the algorithm. In this module this number is chosen mostly by common sense and business criteria. If you're interested in a more technical way of choosing the suitable number and evaluate the stability of your clustering result, you can refer to another documentation [here](http://ethen8181.github.io/machine-learning/clustering/clustering.html).

@@

sessionInfo()

>>>>

---
title: "Finding Groups Within the Data"
author: "Ethen Liu"
date: "2016-04-05"
output:
  rmdformats::readthedown:
    highlight: pygments
---

@@

ggplot( data_sku, aes(CV, ADS) ) + geom_point() +  
geom_hline(yintercept = 4, color = "blue") + 
geom_vline(xintercept = .2, color = "blue") + 
labs( x = "Coefficient of Variation", y = "Average Daily Sales", 
	  title = "SKU Example" )

@@

```{r, results='hide'}

# hierarchical clustering, with k (clustering number) = 3 
d <- dist( scale(data_sku), method = "euclidean" )
cluster <- hclust( d, method = "ward.D" )
data_sku[ ,  groups := as.factor( cutree( cluster, k = 3 ) ) ]

@@

head(data_sku)
ggplot( data_sku, aes(CV, ADS, color = groups) ) + geom_point() + 
labs( x = "Coefficient of Variation", y = "Average Daily Sales", 
	  title = "SKU Example" )

@@

4. For clustering methods, the number of segmentation is a parameter that has to be user-specified before running the algorithm. In this module this number is chosen mostly by common sense and business criteria.

@@

```{r}
devtools::session_info()
```
